---
title: "DATA 606 Data Project Proposal"
author: "Alice Ding"
---

### Data Preparation

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(psych)

# load data
streams <- read.csv("https://raw.githubusercontent.com/addsding/data606/main/project/Streams.csv")
features <- read.csv("https://raw.githubusercontent.com/addsding/data606/main/project/Features.csv")

# append streams data onto features for ease of access
full_data <- merge(x = features, y = streams, by.x = "name", by.y = "Song")
head(full_data)
```


### Research question 

Is there a correlation between tempo and streams (play count) for the top 100 most-played songs on Spotify?

### Cases 

**What are the cases, and how many are there?**

The cases are each songs; the data is comprised of the top 100 most-streamed songs on Spotify.

### Data collection 

**Describe the method of data collection.**

This data was retrieved from [Kaggle](https://www.kaggle.com/datasets/amaanansari09/most-streamed-songs-all-time) by Amaan Ansari. They retrieved it from Spotify's API specifically.

### Type of study 

Observational.



### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

https://www.kaggle.com/datasets/amaanansari09/most-streamed-songs-all-time

### Dependent Variable

**What is the response variable? Is it quantitative or qualitative?**

The response variable would be streams (plays) and it is quantitative. 

### Independent Variable(s)

The explanatory variable would be tempo and it is quantitative. This is the description taken from Spotify:

```The overall estimated tempo of a track in beats per minute (BPM). In musical terminology, tempo is the speed or pace of a given piece and derives directly from the average beat duration.```


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
describe(full_data$Streams..Billions.)
```


```{r}
describe(full_data$tempo)
```
```{r}
ggplot(full_data, aes(x=Streams..Billions.)) + geom_histogram()
```

```{r}
ggplot(full_data, aes(x=tempo)) + geom_histogram()
```