---
title: "Lab1"
author: "Alice Ding"
output: pdf_document
date: "2023-02-01"
---

```{r setup}
library(tidyverse)
library(openintro)
data('arbuthnot', package='openintro')
```

## Exercise 1
### What command would you use to extract just the counts of girls baptized? Try it!


```{r exercise_1}
arbuthnot$girls
```

## Exercise 2
### Is there an apparent trend in the number of girls baptized over the years? How would you describe it? (To ensure that your lab report is comprehensive, be sure to include the code needed to make the plot as well as your written interpretation.)


```{r exercise_2}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_line()
```

There isn't an overarching trend here, just seems to tell the story regarding either birth rates or baptism demand.

It seems like the number of girls baptised was somewhat consistent starting in 1629 before crashing starting in 1640. This would suggest maybe there was a slower birth rate, and this lasted until around 1660 before ticking upwards in a consistent upwards trend until hitting a sort of plateau in the mid 1690s. There were some outlier years (the mid 1660s had a pretty huge crash), but otherwise baptisms were in the 7000s starting in the early 1680s, minus one huge outlier in the early 1700s.

## Exercise 3
### Now, generate a plot of the proportion of boys born over time. What do you see?

```{r new_variable}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)
ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line() +
  geom_smooth(method=lm)
```
I've added a trend line here to show that the proportion of boys being born over time is slowly declining. We see it had an average of ~0.52 to start, but it's been slowly moving towards ~0.515 as time has passed -- it's still above 50% so more boys are being born than girls, but it's almost decreased by 0.01 over the ~80 years for this dataset.